---
title: "Passengers of Okinawa Monorail"
author: "Tohaku"
date: '2023-03-28'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

The Okinawa Monorail is the only public rail in Okinawa, a subtropical resort island. It opened in 2003 and linked 15 stations from Naha Airport to Shuri. It was extended to Tedako-Uranishi in 2019.

Using this Shiny app, you can investigate the number of passengers over time at six important stations. In particular, you can find unusual changes in the following years:

- Sharp drop in 2020 during the pandemic
- Sharp recovery in 2022 after the pandemic

## Sharp drop in 2020 during the pandemic (R expression)

```{r, message=FALSE}
library(dplyr)
library(tidyr)
library(leaflet)
```

```{r}
df <- 
  read.csv("okinawa.csv") |> 
  gather("Station","Passengers",-Year) |> 
  mutate(Station = recode(Station, 
                          "Naha.Airport"="Naha Airport",
                          "Pref.Office"="Prefectural Office",
                          "Tedako.Uranishi"="Tedako-Uranishi"),
         lng = case_when(
           Station == "Naha Airport" ~ 127.65186887791862,
           Station == "Oroku" ~ 127.66723257144604,
           Station == "Prefectural Office" ~ 127.67942052943985,
           Station == "Omoromachi" ~ 127.6979599585009,
           Station == "Shuri" ~ 127.72568327112519,
           Station == "Tedako-Uranishi" ~ 127.74121081707781,
         ),
         lat = case_when(
           Station == "Naha Airport" ~ 26.206456484720434,
           Station == "Oroku" ~ 26.196522024831467,
           Station == "Prefectural Office" ~ 26.21469605242861,
           Station == "Omoromachi" ~ 26.222781042395496, 
           Station == "Shuri" ~ 26.219624113196563,
           Station == "Tedako-Uranishi" ~ 26.24159714008132,
         )) |> 
  group_by(Station) |> 
  mutate(chg = Passengers - lag(Passengers, n=1))
```

```{r}
df <- df |> 
  mutate(radius = abs(chg),
         color = ifelse(chg > 0,"blue","red"))

df_sub <- df |> 
  filter(Year == 2020)
```

```{r,echo=TRUE,eval=FALSE}
leaflet() |>
  addTiles() |> 
  addCircles(data = df_sub,
             lat = ~lat,
             lng = ~lng,
             weight = 1,
             radius = ~sqrt(radius)*10,
             color = ~color,
             popup = ~paste(Station, 
                            "<br>Daily Average:", 
                            format(Passengers,big.mark=","),
                            "<br>Annual Change:", 
                            format(chg,big.mark=",")))
```

## Sharp drop in 2020 during the pandemic

```{r}
leaflet() |>
  addTiles() |> 
  addCircles(data = df_sub,
             lat = ~lat,
             lng = ~lng,
             weight = 1,
             radius = ~sqrt(radius)*10,
             color = ~color,
             popup = ~paste(Station, 
                            "<br>Daily Average:", 
                            format(Passengers,big.mark=","),
                            "<br>Annual Change:", 
                            format(chg,big.mark=",")))
```


## Sharp recovery in 2022 after the pandemic

```{r}
df_sub <- df |> 
  filter(Year == 2022)

leaflet() |>
  addTiles() |> 
  addCircles(data = df_sub,
             lat = ~lat,
             lng = ~lng,
             weight = 1,
             radius = ~sqrt(radius)*10,
             color = ~color,
             popup = ~paste(Station, 
                            "<br>Daily Average:", 
                            format(Passengers,big.mark=","),
                            "<br>Annual Change:", 
                            format(chg,big.mark=",")))
```

